---
title: "Clustering Lab"
author: "Madeleine Ashby"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(plotly)
library(htmltools)
library(devtools)
library(caret)
library(NbClust)
```

Goal: Know how to make decisions and answer questions using clustering. 

Repeat the clustering process only using the Rep house votes dataset. What differences and similarities did you see between how the clustering worked for the datasets?

There were both similarities and differences. Firstly, the most prominent similarity is that both datasets only used 2 centers for clustering.  Another similarity would be that the ratio of the within & between variance accounted for by the 2 clusters was high for both datasets.
The differences were more noticeable. There was a slight difference between the elbow graphs for both datasets in that the Democratic elbow graph showed that 3 clusters may have been a better choice than 2 clusters, while the Republican elbow graph showed that 2 clusters was the optimal choice. Another difference I found between the two datasets was in the first set of plots created.  The Democratic plot appeared to have more overlap - there were blue points in the Republican cluster and there were more outliers.  In the Republican plot, there were far fewer outliers and even fewer miscolored points.  This suggests that party line voting was more present for Republican-introduced bills.


```{r}
#Select the variables to be included in the cluster 
house_votes_Rep = read_csv("C:/Users/Maddie/OneDrive/Desktop/3YEAR/Forked-DS-3001/06_Clustering/house_votes_Rep.csv")
table(house_votes_Rep$party.labels)
View(house_votes_Rep)

# Define the columns to be clustered by sub-setting the data.
# Placing the vector of columns after the comma inside the 
# brackets tells R that you are selecting columns.
clust_data_Rep = house_votes_Rep[, c("aye", "nay", "other")]


```

```{r}
#Run the clustering algo with 2 centers
set.seed(1) #chooses random start location for later comparison
kmeans_obj_Rep = kmeans(clust_data_Rep, centers = 2, 
                        algorithm = "Lloyd")   #<- there are several ways of implementing

```

```{r}
#View the results
kmeans_obj_Rep

#ratio = 79.5%
```

```{r}
#Visualize the output
party_clusters_Rep = as.factor(kmeans_obj_Rep$cluster)


ggplot(house_votes_Rep, aes(x = aye, 
                            y = nay,
                            color = party.labels,  #<- tell R how to color 
                            #   the data points
                            shape = party_clusters_Rep)) + 
  geom_point(size = 6) +
  ggtitle("Aye vs. Nay votes for Republican-introduced bills") +
  xlab("Number of Aye Votes") +
  ylab("Number of Nay Votes") +
  scale_shape_manual(name = "Cluster", 
                     labels = c("Cluster 1", "Cluster 2"),
                     values = c("1", "2")) +
  scale_color_manual(name = "Party",         #<- tell R which colors to use and
                     #   which labels to include in the legend
                     labels = c("Democratic", "Republican"),
                     values = c("blue", "red")) +
  theme_light()
```

```{r}
#Evaluate the quality of the clustering 

# Inter-cluster variance,
# "betweenss" is the sum of the distances between points 
# from different clusters.
num_Rep = kmeans_obj_Rep$betweenss

# Total variance, "totss" is the sum of the distances
# between all the points in the data set.
denom_Rep = kmeans_obj_Rep$totss

# Variance accounted for by clusters.
(var_exp_Rep = num_Rep / denom_Rep)

```

```{r}
#Use the function we created to evaluate several different number of clusters
explained_variance = function(data_in, k){
  
  # Running the kmeans algorithm.
  set.seed(1)
  kmeans_obj = kmeans(data_in, centers = k, algorithm = "Lloyd", iter.max = 30)
  
  # Variance accounted for by clusters:
  # var_exp = intercluster variance / total variance
  var_exp = kmeans_obj$betweenss / kmeans_obj$totss
  var_exp  
}

explained_var_Rep = sapply(1:10, explained_variance, data_in = clust_data_Rep)
explained_var_Rep

```

```{r}
#Create a elbow chart of the output 
elbow_data_Rep = data.frame(k = 1:10, explained_var_Rep)

ggplot(elbow_data_Rep, 
       aes(x = k,  
           y = explained_var_Rep)) + 
  geom_point(size = 4) +           #<- sets the size of the data points
  geom_line(size = 1) +            #<- sets the thickness of the line
  xlab('k') + 
  ylab('Inter-cluster Variance / Total Variance') + 
  theme_light()

#looks like 2 clusters is best based on elbow plot

```

```{r}
#Use NbClust to select a number of clusters
(nbclust_obj_Rep = NbClust(data = clust_data_Rep, method = "kmeans"))

# View the output of NbClust.
nbclust_obj_Rep

# View the output that shows the number of clusters each method recommends.
View(nbclust_obj_Rep$Best.nc)

```

```{r}
#Display the results visually 
freq_k_Rep = nbclust_obj_Rep$Best.nc[1,]
freq_k_Rep = data.frame(freq_k_Rep)
View(freq_k_Rep)

# Check the maximum number of clusters suggested.
max(freq_k_Rep)

#essentially resets the plot viewer back to default
#dev.off()

# histogram plot
ggplot(freq_k_Rep,
       aes(x = freq_k_Rep)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 15, by = 1)) +
  scale_y_continuous(breaks = seq(0, 12, by = 1)) +
  labs(x = "Number of Clusters",
       y = "Number of Votes",
       title = "Cluster Analysis")


```

```{r}
#Using the recommended number of cluster compare the quality of the model 
#with 2 clusters 

#The recommended number of clusters using the elbow graph and nbclust was 2, so the two models will be equivalent.


```


```{r}
#Bonus: Create a 3d version of the output
party_color3D_Rep = data.frame(party.labels = c("Democrat", "Republican"),
                               color = c("blue", "red"))

#View(party_color3D_Rep)


# (inner) Join the new data frame to our house_votes_Dem data set.
# when you inner join, they do not need the same dimensions, but you need the same column name!
house_votes_color_Rep = inner_join(house_votes_Rep, party_color3D_Rep)

#adds the cluster column
house_votes_color_Rep$clusters <- (party_clusters_Rep)

#View(house_votes_color_Rep)

#removes characters that aren't going to be parseable
house_votes_color_Rep$Last.Name <- gsub("[^[:alnum:]]", "", house_votes_color_Rep$Last.Name)

# Use plotly to do a 3d imaging 

fig <- plot_ly(house_votes_color_Rep, 
               type = "scatter3d",
               mode="markers",
               symbol = ~clusters,
               x = ~aye, 
               y = ~nay, 
               z = ~other,
               color = ~color, # ~ means "identify just this variable and use all layers (plotly)
               colors = c('#0C4B8E','#BF382A'), 
               text = ~paste('Representative:',Last.Name,
                             "Party:",party.labels))


fig

```

In a separate Rmarkdown document work through a similar process 
with the NBA data (nba2020-21 and nba_salaries_21), merge them together. 

You are a scout for the worst team in the NBA, probably the Wizards. Your general manager just heard about Data Science and thinks it can solve all the teams problems!!! She wants you to figure out a way to find players that are high performing but maybe not highly paid that you can steal to get the team to the playoffs! 

Details: 

- Determine a way to use clustering to estimate based on performance if 
players are under or over paid, generally. 
- Then select three players you believe would be best your team and explain why. 
- Provide a well commented and clean (knitted) report of your findings that can 
be presented to your GM. Include a rationale for variable selection, details 
on your approach and a overview of the results with supporting visualizations. 
- Create a new repo for this assignment either for yourself and 
submit the link along with your knitted report. 

Hints:

- Salary is the variable you are trying to understand 
  - don't include it in your clustering process though
    - figure out if the other performance variables cluster, independent of salary
    - use TWO variables that are highly correlated with salary
        - party affiliation is the equivalent here
            - we used colors to represent this, so we want to color by salary!
            
- You can include numerous performance variables in the clustering but when 
interpreting you might want to use graphs that include variables that are the 
most correlated with Salary
- You'll need to standardize the variables before performing the clustering
- Be specific about why you selected the players that you did, more detail is 
better
- Use good coding practices, comment heavily, indent, don't use for loops unless
totally necessary and create modular sections that align with some outcome. If 
necessary create more than one script,list/load libraries at the top and don't 
include libraries that aren't used.

submit : original Rmd and html files for both!




