---
title: "Clustering Lab"
author: "Brian Wright"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Repeat the clustering process only using the Rep house votes dataset
- What differences and similarities did you see between how the clustering 
worked for the datasets?

```{r}
#Select the variables to be included in the cluster 

```

```{r}
#Run the clustering algo with 2 centers
```

```{r}
#View the results
```

```{r}
#Visualize the output
```

```{r}
#Create a 3d version of the output
```

```{r}
#Evaluate the quality of the clustering 
```

```{r}
#Use the function we created to evaluate several different number of clusters
```

```{r}
#Create a elbow chart of the output 
```

```{r}
#Use NbClust to select a number of clusters
```

```{r}
#Display the results visually 
```

```{r}
#Using the recommended number of cluster compare the quality of the model 
#with 2 clusters 

-3/4 * log2(3/4) -1/4* log2(1/4)

-5/14*log2(5/14)-4/14*log2(4/14)-5/14*log2(5/14)

((3/4) * 0.92) + ((1/4) * 0)
.5*1
```

In a separate Rmarkdown document work through a similar process 
with the NBA data. 

You are a scout for the worst team in the NBA, probably the Wizards. Your 
general manager just heard about Data Science and thinks it can solve all the
teams problems!!! She wants you to figure out a way to find players that are 
high performing but maybe not highly paid that you can steal to get your team 
out of the toilet! 

Details: 

- Determine a way to use clustering to estimate based on performance if 
players are under or over paid, generally.  
- Provide a well commented and clean (knitted) report of your findings that can 
be presented to your GM. Include a rationale for variable selection, details 
on your approach and a overview of the results with supporting visualizations.  

Hints:

- Salary is the variable you are trying to understand 
- You can include numerous performance variables in the clustering but when 
interpreting you might want to use graphs that include variables that are the 
most correlated with Salary, also might be a good way to select variables to
include
- You'll need to standardize the variables before performing the clustering 
  





