---
title: "NBA Clustering"
author: "Madeleine Ashby"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(plotly)
library(htmltools)
library(devtools)
library(caret)
library(NbClust)
```

```{r}
#Read in data

nba1 <- read.csv("C:/Users/Maddie/OneDrive/Desktop/3YEAR/Forked-DS-3001/data/nba_salaries_21.csv")
nba2 <- read.csv("C:/Users/Maddie/OneDrive/Desktop/3YEAR/Forked-DS-3001/data/nba2020-21.csv")

nba <- merge(nba1, nba2)
nba <- na.exclude(nba)

```

Variable selection:
In order to determine which players are high performing but not as highly paid, I will be using clustering to understand the salary variable.  My variables of choice here will be the minutes played, threes scored and total points scored, as I believe that a good player will play more minutes, score more threes, and have an overall greater points scored total.


```{r}
#Normalize data

normalize <- function(x){
 (x - min(x)) / (max(x) - min(x))
}

nba$minutes_played <- normalize(nba$MP)
nba$threes <- normalize(nba$X3PA)
nba$points <- normalize(nba$PTS)

#cut salary
#summary(nba$X2020.21)
#nba$salary <- cut(nba$X2020.21, c(0, 2113500,  4568480, 11992265, 43006362),labels = c("Lowest", "Low", "High", "Highest"))
nba$salary <- nba$X2020.21

#Select clustering data!
clust_nba = nba[, c(31:33)]
View(clust_nba)

```

```{r}
#Run the clustering algorithm with 2 centers
set.seed(1)
kmeans_obj = kmeans(clust_nba, centers = 2, 
                        algorithm = "Lloyd")

kmeans_obj
# we can see here that the ratio with 2 centers is only 59.6%... not so great.

#Determine optimal number of clusters

#Method 1: Elbow Graph
explained_variance = function(data_in, k){
  set.seed(1)
  kmeans_obj = kmeans(data_in, centers = k, algorithm = "Lloyd", iter.max = 30)
  var_exp = kmeans_obj$betweenss / kmeans_obj$totss
  var_exp  
}

explained_var = sapply(1:10, explained_variance, data_in = clust_nba)
explained_var

elbow_data = data.frame(k = 1:10, explained_var)

ggplot(elbow_data, 
       aes(x = k,  
           y = explained_var)) + 
  geom_point(size = 4) +           #<- sets the size of the data points
  geom_line(size = 1) +            #<- sets the thickness of the line
  xlab('k') + 
  ylab('Inter-cluster Variance / Total Variance') + 
  theme_light()

#Method 2: Nbclust
(nbclust_obj = NbClust(data = clust_nba, method = "kmeans"))
nbclust_obj
View(nbclust_obj$Best.nc)

#Both methods show that 3 is the optimal number of clusters!

```

```{r}
#Run algorithm agian with 3 centers.
set.seed(1)
kmeans_obj = kmeans(clust_nba, centers = 3, 
                        algorithm = "Lloyd")

kmeans_obj

#now the ratio is 76.4% - much better.
```

```{r}
#Visualizing output
salary_clusters <- as.factor(kmeans_obj$cluster)
#salary_clusters

ggplot(nba, aes(x = minutes_played, 
                            y = threes,
                            color = salary,  #<- tell R how to color 
                            #   the data points
                            shape = salary_clusters)) + 
  geom_point(size = 6) +
  ggtitle("Threes Scored vs Minutes Played for NBA Players") +
  xlab("Number of Minutes Played") +
  ylab("Number of Threes Scored") +
  scale_shape_manual(name = "Cluster", 
                     labels = c("Cluster 1", "Cluster 2", "Cluster 3"),
                     values = c("1", "2", "3")) +
  #scale_color_manual(name = "Salary",         #<- tell R which colors to use and
                     #   which labels to include in the legend
                    # labels = c("Lowest", "Low", "High", "Highest"),
                     #values = c("red", "orange", "yellow", "green")) +
  theme_light()
```

```{r}
#Bonus: Create a 3d version of the output
#no need to join datasets/define colors for different salaries 
#because we are treating salary as a continuous variable!
#color will be a gradient

#adds the cluster column
nba$clusters <- (salary_clusters)

#removes characters that aren't going to be parseable
nba$Player <- gsub("[^[:alnum:]]", "", nba$Player)

# Use plotly to do a 3d imaging 
fig <- plot_ly(nba, 
               type = "scatter3d",
               mode="markers",
               symbol = ~clusters,
               x = ~minutes_played, 
               y = ~threes, 
               z = ~points,
               color = ~salary, # ~ means "identify just this variable and use all layers (plotly)
               text = ~paste('Player: ',Player,
                             "Salary: ", salary))

fig
```
In this document I have analyzed salary and performance measures of different NBA players to ultimately determine whether players are generally over- or under-paid.  After close examination using k-means clustering with the variables of minutes played, threes scored, overall points scored, and salary, it is clear that NBA players are generally underpaid.  This is clearly shown in the first plot, as we can see that there are not many light blue points on towards the origin, but there are several darker points as the graph approaches infinity.

Using the 3d plot, I was able to identify three players I believe would be best to add to our team. The following players would be good additions to our team because they perform well but are currently being paid low salaries.  Thus, it would not be difficult to offer them a better salary and improve our team.  The players are: Donovan Mitchell, DeAaron Fox, and Trae Young.