<<<<<<< HEAD
---
title: "Reinforcement Lab"
author: "Courtney Kennedy"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(NbClust)
library(ggpubr)
```

```{r}
game_df <- read_csv("data-frame.csv")
```
```{r}
game_df %>%
  filter(num_turns < 10) %>%
  ggplot(aes(x = cora, y = corc, color = won)) +
  geom_point()
```
```{r}
cluster_data <- game_df[,c(4,6,7)]
normalize <- function(x){
 (x - min(x)) / (max(x) - min(x))
}
cluster_data$num_turns <- normalize(cluster_data$num_turns)

kmeans_obj = kmeans(cluster_data, centers = 5, algorithm = "Lloyd")

clusters_game = as.factor(kmeans_obj$cluster)

game_df %>%
  ggplot(aes(x = cora, y = corc, color = clusters_game)) + 
  geom_point(size = 6) +
  theme_light()
```

```{r}

# Run NbClust.
nbclust_obj <-  NbClust(data = cluster_data, method = "kmeans")


# View the output that shows the number of clusters each method recommends.
View(nbclust_obj$Best.nc)

freq_k = nbclust_obj$Best.nc[1,]
freq_k = data.frame(freq_k)


# Plot as a histogram.
ggplot(freq_k,
       aes(x = freq_k)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 15, by = 1)) +
  scale_y_continuous(breaks = seq(0, 12, by = 1)) +
  labs(x = "Number of Clusters",
       y = "Number of Votes",
       title = "Cluster Analysis")
```

```{r}
kmeans_obj = kmeans(cluster_data, centers = 3, algorithm = "Lloyd")
clusters_game = as.factor(kmeans_obj$cluster)
game_df %>%
  ggplot(aes(x = cora, y = corc, color = clusters_game)) + 
  geom_point(size = 6) +
  theme_light()
```
```{r}
game_df %>%
  ggplot(aes(x = cora, y = num_turns, color = clusters_game)) + 
  geom_point(size = 6) +
  theme_light()
```
```{r}
kmeans_obj = kmeans(cluster_data, centers = 5, algorithm = "Lloyd")
clusters_game = as.factor(kmeans_obj$cluster)

game_df_cluster_color <- tibble(game_df, cluster = clusters_game)

win_cluster <- game_df_cluster_color %>%
  filter(won == TRUE) %>%
  ggplot(aes(x = cora, y = corc, color = cluster)) + 
  geom_point(size = 6) +
  theme_light() +
  ggtitle("Won Game")

lose_cluster <- game_df_cluster_color %>%
  filter(won == FALSE) %>%
  ggplot(aes(x = cora, y = corc, color = cluster)) + 
  geom_point(size = 6) +
  theme_light() +
  ggtitle("Lost Game")


ggarrange(win_cluster, lose_cluster)

```

=======
---
title: "Reinforcement Lab"
author: "Courtney Kennedy"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(NbClust)
library(ggpubr)
```

```{r}
game_df <- read_csv("data-frame.csv")
```
```{r}
game_df %>%
  filter(num_turns < 10) %>%
  ggplot(aes(x = cora, y = corc, color = won)) +
  geom_point()
```
```{r}
cluster_data <- game_df[,c(4,6,7)]
normalize <- function(x){
 (x - min(x)) / (max(x) - min(x))
}
cluster_data$num_turns <- normalize(cluster_data$num_turns)

kmeans_obj = kmeans(cluster_data, centers = 5, algorithm = "Lloyd")

clusters_game = as.factor(kmeans_obj$cluster)

game_df %>%
  ggplot(aes(x = cora, y = corc, color = clusters_game)) + 
  geom_point(size = 6) +
  theme_light()
```

```{r}

# Run NbClust.
nbclust_obj <-  NbClust(data = cluster_data, method = "kmeans")


# View the output that shows the number of clusters each method recommends.
View(nbclust_obj$Best.nc)

freq_k = nbclust_obj$Best.nc[1,]
freq_k = data.frame(freq_k)


# Plot as a histogram.
ggplot(freq_k,
       aes(x = freq_k)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 15, by = 1)) +
  scale_y_continuous(breaks = seq(0, 12, by = 1)) +
  labs(x = "Number of Clusters",
       y = "Number of Votes",
       title = "Cluster Analysis")
```

```{r}
kmeans_obj = kmeans(cluster_data, centers = 3, algorithm = "Lloyd")
clusters_game = as.factor(kmeans_obj$cluster)
game_df %>%
  ggplot(aes(x = cora, y = corc, color = clusters_game)) + 
  geom_point(size = 6) +
  theme_light()
```
```{r}
game_df %>%
  ggplot(aes(x = cora, y = num_turns, color = clusters_game)) + 
  geom_point(size = 6) +
  theme_light()
```
```{r}
kmeans_obj = kmeans(cluster_data, centers = 5, algorithm = "Lloyd")
clusters_game = as.factor(kmeans_obj$cluster)

game_df_cluster_color <- tibble(game_df, cluster = clusters_game)

win_cluster <- game_df_cluster_color %>%
  filter(won == TRUE) %>%
  ggplot(aes(x = cora, y = corc, color = cluster)) + 
  geom_point(size = 6) +
  theme_light() +
  ggtitle("Won Game")

lose_cluster <- game_df_cluster_color %>%
  filter(won == FALSE) %>%
  ggplot(aes(x = cora, y = corc, color = cluster)) + 
  geom_point(size = 6) +
  theme_light() +
  ggtitle("Lost Game")


ggarrange(win_cluster, lose_cluster)

```

>>>>>>> 0727dce31d03fcbb9d8925e72c76339909e271ca
