---
title: "reinforcement"
author: "Teagan Norrgard, trn8cwf"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### the purpose of this code was to practice some clustering and practice visualization through the ggplot package
### after this, I looked up different ways to add color and different aesthetics to the plot, since I did not get much practice with color in this

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)


# read in data and create dataframe (df1)
df <- read_csv("/Users/teagannorrgard/DS-3001/week-11-reinforcement-lab/data-summary.csv")
df1 <- select(df,main_colors,opp_colors,on_play,num_turns,won)

# feature engineering (cora,corc)
df2 <- select(df,"deck_Adeline, Resplendent Cathar":"deck_Wrenn and Seven")
mat = data.matrix(df2)
vec1 <- vector()
vec3 <- vector()
for(i in 1:nrow(mat) ){
  x<-cor( mat[1,] , mat[i,])
  vec1 <- c(vec1,x)
  z<-cor( mat[47,] , mat[i,])
  vec3 <- c(vec3,z)
}

# add new features to dataframe
df1 <- df1 %>% mutate(cora = vec1)
df1 <- df1 %>% mutate(corc = vec3)

# make scatter plot comparing new features
ggplot(df1,aes(x=cora,y=corc))+geom_point()
```


```{r}
library(NbClust)

clust_data = df1[, c("corc", "cora")]

# Run NbClust.
(nbclust_obj = NbClust(data = clust_data, method = "kmeans"))

## this gave me a plot that suggested 5 was the optimal amount of clusters

set.seed(1)
kmeans_obj = kmeans(clust_data, centers = 5, 
                        algorithm = "Lloyd")

clusters = as.factor(kmeans_obj$cluster)
```

```{r}
ggplot(clust_data, aes(x = cora, y = corc, shape = clusters) + 
         geom_point(size = 6) +
         ggtitle("corA vs corC") +
         xlab("corA") +
         ylab("corC") +
         scale_shape_manual(name = "Cluster",
                            labels = c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5"),
                            values = c("1", "2", "3", "4", "5")) + theme_light())
```
my group and I cannot figure out what is going one with my plot and why it will not run. We all started with the same basic cluster algorithm and used the same functions to form our outline of the plot. We then were going to go our separate ways trying to make different visualizations with our cluster model. None of us can figure out why my plot will not run.

but in trying to figure out this error I definitely learned a lot about ggplot through research of my error, and what arguments I can use and how I can manipulate my plots (even though I couldn't see my results :( )


