---
title: "dplyr_Lab_Assignment"
author: "Brian Wright"
date: "2/9/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)


```


################## AND NOW THE EXCITEMENT BEGINS! DPLYR AND TIDYVERSE PRACTICE 

## As a reminder five key dplyr functions that allow you to solve "the vast majority of your data-manipulation
## challenges" are:

## 1) filter() -- Pick observations by their values
## 2) arrage() -- Reorder the rows
## 3) select() -- Pick variables by their names
## 4) mutate() -- Create new variables with functions of existing variables
## 5) summarize() -- Collapse many values down to a single summary
## Bonus: group_by -- we can use this in combination with any of the five to change the scope of
## each function from operating on the entire dataset to operating on it 
group-by-group.


For this assignment you can use any or all of the datasets I 
provided or choose two of your own!  I would like you to compose 
two question for each dataset (or dataset of your choosing) that requires at a minimum three functions listed above to answer. Do not repeat function sequences or questions.  Information or the actual data for the datasets can be found through the links below. In summary for each dataset produce four parts: 1)Question, 2)psuedocode that answers the question, 3)single line transactions using dplyr functions, 4) pipe the individual lines together and create a function. 

Also you can work with your groups for coding tips/advise or work through similar programming issues but everyone must have their own set of questions and results. 

# any dataset in the data file is fine or 

# nf2008_fg: http://users.stat.ufl.edu/~winner/data/nfl2008_fga.csv

# red_wine_quality: https://data.world/uci/wine-quality


# Lets walk through a example. Below is a dataset on students admitted to a graduate program in Data Science (actually don't remember where this data came from). 

## Are there differences in the admitted number of students on average by GPA and school rank?

  >A easy way to handle this is to identify function calls starting at the end of the question and working backwards. 

## Pseudocode 
  - Group by school rank
  - Calculate GPA Average
  - Count total admitted 
  - Calculate % Admitted
  - Arrange in descending order by Mean GPA


Loading and preparing the data 
```{r}
admit_2 <- read_csv("data/LogReg.csv")

str(admit_2)#str is short for structure and provides a overview of the data
#What do you see here?

#creating a quick list to use as a index
names <- c("admit","rank")

names
#changing those two variables to factors, lapply is a function that takes a function and applies the given function to your data

admit_2[,names] <- lapply(admit_2[,names],factor)#In this example our data is "admit_2" we are using "[ ,names]" to only target the admit and rank variables and the function we are placing into lapply is "factor" which changes given variables to factors. 

str(admit_2)

```

Single line version 
```{r}

View(admit_2) # Taking a look at the data
x1 <- group_by(admit_2,rank)# grouping by the factor variables, done in the background 
# Calculate additional variables
View(x1)
x2 <- summarize(x1,admit=sum(admit == 1),count=n(),mean(gpa))# using x1 from above we are generating three variables a 
  # sum of all 1s in the admit column for each college rank,
  # total count of in each category of rank 
  # mean GPA in each of the rankings
View(x2)
x2$ave_admit <- x2$admit/x2$count # added a variable ave to the x2,by dividing total admitted by total count for each of the ranking levels. The "$" is a atomic symbol that's used to local specific columns in our dataframe

x2

x3 <- arrange(x2,desc(`mean(gpa)`))# Arrange by rank descending 
View(x3)
## Are there differences in the admitted number of students on average by GPA and school rank?

```

## Piping it all together. "%>%" in tidyverse simply take the information from the previous row of code and carry it over to the next line. 
```{r}
#creating a dplyr version
(admit_1 <- group_by(admit_2, rank)%>%
 summarize(admit=sum(admit == 1),count=n(),mean(gpa))%>%# uses functions to create new variables
 mutate(ave_admit=admit/count)%>%#creating the average using the variables from above. 
 arrange(desc(`rank`)))
# extra () at beginning and end tell R to print the result, the %>% sign is used for dplyr functions and essentially takes the results from the previous line and inputs into the next. 

```


## Now just turn it into a function, so if you needed to repeat this data request on future datasets it would be simple. 
```{r}
ave_admit <- function(x){
  xx <- group_by(x, rank)%>%
  summarize(admit=sum(admit == 1),count=n(),mean(gpa))%>%
  mutate(ave_admit=admit/count)%>%  
  arrange(desc(`mean(gpa)`))
  print(xx)
}#so this function has one attribute x, which will be the data we pass into the function once it's created. 
```

```{r}
ave_admit(admit_2)# the above code creates the function, this code runs the function. The results of which are the same.

#What do about these results are interesting? 

```
