---
title: "Decision Trees"
author: "Brian Wright"
date: "November 27, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Libraries
```{r}
library(rio)
library(plyr)
library(tidyverse)
library(rpart)
library(psych)
library(pROC)
#install.packages("rpart.plot")
library(rpart.plot)
#install.packages("rattle")
library(rattle)
library(caret)
library(C50) #Need this to pass into caret 
library(mlbench)
```


Let's take a look at the caret method using rpart, nice overview of 

CART Example using rpart: Use a new Dataset binary and doing three data
partitions: Training, Tuning and Testing
```{r}

winequality <- read.csv("data/winequality-red-ddl.csv")
View(winequality)
str(winequality)

#drop 12 predicts text_rank perfectly

winequality <- winequality[,-12]

sum(is.na(winequality))#got some NAs to deal with here...what should we do?
```


## Missing Data 
```{r}
summary(winequality)#show the location by variable

mice::md.pattern(winequality)#provides information on the location of the NAs

winequality <- winequality[complete.cases(winequality),]

str(winequality)
```

## Collapsing the target
```{r}
table(winequality$text_rank)


winequality$text_rank <- fct_collapse(winequality$text_rank,
                                      ave=c("ave","average-ish","poor","poor-ish",""),
                                      excellent = c("excellent","good"))

#Prevalance
203/(1409+203)
```


## Splitting the Data: 
```{r}
#There is not a easy way to create 3 partitions using the createDataPartitions
#so we are going to use it twice. Mostly because we want to stratify on the variable we are working to predict. What does that mean?  

set.seed(1999)
part_index_1 <- caret::createDataPartition(winequality$text_rank,
                                           times=1,
                                           p = 0.70,
                                           groups=1,
                                           list=FALSE)

train <- winequality[part_index_1, ]
tune_and_test <- winequality[-part_index_1, ]
train
#The we need to use the function again to create the tuning set 

tune_and_test_index <- createDataPartition(tune_and_test$text_rank,
                                           p = .5,
                                           list = FALSE,
                                           times = 1)

tune <- tune_and_test[tune_and_test_index, ]
test <- tune_and_test[-tune_and_test_index, ]


dim(train)
dim(test) 
dim(tune)

```


```{r}
# Choose the features and classes, slightly different approach for caret, need to create features and target sets from the training data.

features <- train[,-12]#dropping 12 because it's target variable. 
View(features)
target <- train$text_rank

str(features)
str(target)

#Three steps in building a caret ML model
#Step 1: Cross validation process-the process by which the training data will be used to build the initial model must be set. As seen below:

?trainControl

fitControl <- trainControl(method = "repeatedcv",
                          number = 10,
                          repeats = 5,
                          classProbs = TRUE,
                          summaryFunction = twoClassSummary) 
# number - number of folds
# repeats - number of times the CV is repeated, takes the average of these repeat rounds
# ClassProbs - Yes or no,indicating if we should include the class probabilities 
# summaryFunction - different options for producing evaluation metrics. #review the documentation on https://topepo.github.io/caret/measuring-performance.html


#Step 2: Usually involves setting a hyper-parameter search. This is optional and the hyper-parameters vary by model. Let's take a look at the documentation for the model we are going to use. 

tree.grid <- expand.grid(maxdepth=c(5,7,9,11))
#let's look at the documentation in two places 
# for the tune grid function: https://topepo.github.io/caret/model-training-and-tuning.html

#options for the rpart2: https://topepo.github.io/caret/train-models-by-tag.html#tree-based-model

#Step 3: Train the models
set.seed(1984)
wine_mdl <- train(x=features,
                y=target,
                method="rpart2",
                trControl=fitControl,
                metric="ROC")


set.seed(1984)
wine_mdl_1 <- train(x=features,
                y=target,
                method="rpart2",#type of model uses maxdepth to select a model
                trControl=fitControl,#previously created
                tuneGrid=tree.grid,#expanded grid
                metric="ROC")#selected on of the metrics available from two variable summary.


```


## Let's see how we did. 
```{r}
wine_mdl
wine_mdl_1

plot(wine_mdl)
varImp(wine_mdl)

rpart.plot(wine_mdl$finalModel, type=4,extra=101)

wine_mdl$results

```


Let's use the model to predict and the evaluate the performance
```{r}
#build a function to make this a little easier 

predictandCM<- function(model,data,modeltype,ref)
{
  #model using, data going into the model, and output type for predict function
  pred <-predict(model,data,type=modeltype)
  confusionMatrix(pred, reference=ref, positive = 'excellent')
}


predictandCM(wine_mdl,tune,"raw",tune$text_rank)#compare to the alternative model

predictandCM(wine_mdl_1,tune,"raw",tune$text_rank)#looks like the orginal model is likely better 

#Can also do this without the function
wine_pred_tune = predict(wine_mdl,tune,tune$text_rank, type= "prob")#probabilities

wine_pred_tune_labels = predict(wine_mdl,tune,tune$text_rank,type = "raw")#Labels

View(as_tibble(wine_pred_tune_labels))

#Lets use the confusion matrix
wine_eval <- caret::confusionMatrix(wine_pred_tune_labels, 
                as.factor(tune$text_rank), 
                dnn=c("Prediction", "Actual"),
                positive="excellent",
                mode = "everything")

wine_eval
```


## Let's play with the threshold, what do we see 
```{r}
plot(density(wine_pred_tune$excellent))

adjust_thres <- function(x, y, z) {
  #x=pred_probablities, y=threshold, z=tune_outcome
  thres <- as.factor(ifelse(x > y, "excellent","ave"))
  confusionMatrix(thres, z, positive = "excellent", dnn=c("Prediction", "Actual"), mode = "everything")
}

adjust_thres(wine_pred_tune$excellent,y=.4,tune$text_rank)

```



## Feature Engineering: 
quick aside, check out [Feature Engineering for ML](https://learning.oreilly.com/library/view/feature-engineering-for/9781491953235/ch02.html#idm140610104790176). It's a very good introduction
to this topic and free with your O'Reilly subscription through UVA. 

```{r}
summary(winequality)

hist(features$total.sulfur.dioxide, plot = TRUE)

fivenum((winequality$total.sulfur.dioxide))

total.sulfur.dioxide <- cut(winequality$total.sulfur.dioxide,c(0,37,300), labels = c("low","high"))


table(total.sulfur.dioxide)

winequality_1 <- data.frame(winequality[,-7],total.sulfur.dioxide)

#data 

set.seed(1999)
part_index_1 <- caret::createDataPartition(winequality_1$text_rank,
                                           times=1,
                                           p = 0.70,
                                           groups=1,
                                           list=FALSE)

train_1 <- winequality_1[part_index_1, ]
tune_and_test_1 <- winequality_1[-part_index_1, ]

#The we need to use the function again to create the tuning set 

tune_and_test_index <- createDataPartition(tune_and_test_1$text_rank,
                                           p = .5,
                                           list = FALSE,
                                           times = 1)

tune_1 <- tune_and_test_1[tune_and_test_index, ]
test_1 <- tune_and_test_1[-tune_and_test_index, ]

str(train_1)

features_3 <- train_1[,-11]
target_1 <- train_1$text_rank

View(features_3)

set.seed(1984)
wine_mdl_2 <- train(x=features_3,
                y=target_1,
                method="rpart2",
                trControl=fitControl,
                tuneGrid=tree.grid,
                metric="ROC")

wine_mdl_2



predictandCM(wine_mdl,tune,"raw",tune$text_rank)
predictandCM(wine_mdl_1,tune,"raw",tune$text_rank)
predictandCM(wine_mdl_2,tune_1,"raw",tune_1$text_rank)

```

## We can also review the variable importance

```{r}
varImp(wine_mdl)
varImp(wine_mdl_1)
varImp(wine_mdl_2)

```

## Predict with test, how did we do? 
```{r}
predictandCM(wine_mdl,test,"raw",test$text_rank)
predictandCM(wine_mdl_1,test,"raw",test$text_rank)
predictandCM(wine_mdl_2,test,"raw",test$text_rank)
```


Another example using just the rpart without the caret wrapper
```{r}
tree_example <- tibble(import("data/pregnancy.csv", check.names= TRUE))

describe(tree_example)
View(tree_example)
str(tree_example)

#We want to build a classifier that can predict whether a shopper is pregnant based on the items they buy so we can direct-market to that customer if possible. 


sum(tree_example$PREGNANT)
length(tree_example$PREGNANT)

(x <- 1- sum(tree_example$PREGNANT)/length(tree_example$PREGNANT))


#What does .72 represent in this context? 

```

#reformat for exploration purposes
```{r}

#Creating a vertical dataframe for the pregnant variable, just stacking the variables on top of each other. 


tree_example_long = tree_example %>% gather(Var, #<- list of predictor variables
                                Value,#<- the values of those predictor variables
                                -PREGNANT)  #<- the column to gather the data by

View(tree_example)
View(tree_example_long)

```


#See what the base rate of likihood of pregnancy looks like for each variable
```{r}
# Calculate the probability of being pregnant by predictor variable.
# Since the data is binary you can take the average to get the probability.

#Older way, but works well for doing multi-level group summaries, creates new variables for each group versus a summary for the entire list. 

tree_example_long_form = ddply(tree_example_long, 
                            .(Var, Value),#<- group by Var and Value, "." allows us to call the variables without quoting
                            summarize,  
                            prob_pregnant = mean(PREGNANT), #<- probability of being pregnant
                            prob_not_pregnant = 1 - mean(PREGNANT)) #<- probability of not being pregnant

#?ddply

View(tree_example_long_form)
```

#Build  the model using rpart and CART Algo (Gini Index - Binary Tree)
```{r}
# In order for this decision tree algorithm to run, 
# all the variables will need to be turned into factors. 
#Make sure your variables are classified correctly. 


tree_example = lapply(tree_example, function(x) as.factor(x))

#This is a handy reference on apply(), lapply(), sapply() are 
#all essentially designed to avoid for loops, especially in combination 
#with (function (x))

#https://www.r-bloggers.com/using-apply-sapply-lapply-in-r/

str(tree_example)

tree_example <- as_tibble(tree_example)

table(tree_example$PREGNANT)

#Also want to add data labels to the target
tree_example$PREGNANT <- factor(tree_example$PREGNANT,labels = c("not_preg", "preg"))

#Build the model
# Train the tree with the rpart() function.
# We'll need to set the seed to make the results reproducible. 
set.seed(1980)
tree_example_tree_gini = rpart(PREGNANT~.,  #<- formula, response variable ~ predictors
                           #   "." means "use all other variables in data"
                            method = "class",#<- specify method, use "class" for tree
                            parms = list(split = "gini"),#<- method for choosing tree split
                            data = tree_example,#<- data used
                            control = rpart.control(cp=.001))

#Look at the results
tree_example_tree_gini

View(tree_example_tree_gini$frame)

# dev - the deviance or the total sum of squares within the node, so if
#       you divide this by the sample size in each node you get the variance
# yval - average value of the trait at the node (for categorical values identifies the group)  
# complexity - the value of the parameter used to make the split (gini or information gain)
# ncompete - number of competing variables that can be considered for this split
# nsurrogate - number of surrogate trees (used when there is missing data in the test data set, to mimic the effects of splits in the training data set)
# yval2 - average value of the trait at the node (for categorical values identifies the group), although it can mean different things when the rpart function is used for regression trees or other analyses 


rpart.plot(tree_example_tree_gini, type =4, extra = 101)#package rpart.plot
#export this to  pdf for better viewing
?rpart.plot

#The "cptable" element includes the optimal prunnings based on the complexity parameter.

View(tree_example_tree_gini$cptable)

plotcp(tree_example_tree_gini)#Produces a "elbow chart" for various cp values

# Here's a summary:
# CP - complexity parameter, or the value of the splitting criterion (gini or information gain)
# nsplit - number of splits
# rel error - the relative error rate for predictions for the data that generated the tree
# xerror - cross-validated error, default cross-validation setting uses 10 folds
# xstd - the standard derivation of cross-validated errors

# NOTE: 
# For pruning a tree, the rule of thumb is to choose the split at the lowest level 
# where the rel_error + xstd < xerror

cptable_ex <- as_tibble(tree_example_tree_gini$cptable, )
str(cptable_ex)

cptable_ex$opt <- cptable_ex$`rel error`+ cptable_ex$xstd

View(cptable_ex)

# Ok so let's compare the cptable_ex, the cpplot and the decision tree plot, 
# they all covered around 8ish splits of the tree or a cp of .014ish. 
# Print out skips splits that result in terminal leaf nodes for 
# some reason, so makes it a little hard to interpret 

rpart.plot(tree_example_tree_gini, type =4, extra = 101)

# Shows the reduction in error provided by including a given variable 
tree_example_tree_gini$variable.importance

```

#Plot the Output to png 

```{r}
# Plot tree, and save to a png file.
png("Pregnancy_tree_gini.png",  #<- image name
    width = 1000,               #<- width of image in pixels
    height = 600)               #<- height of image in pixels

post(tree_example_tree_gini,                  #<- the rpart model to plot
     file = "",                            #<- ensure the png file is created correctly
     title = "Tree for Pregnancy - gini")  #<- the title of the graph

dev.off()
```


# Test the accuracy 
```{r}
# Let's use the "predict" function to test our our model and then 
# evaluate the accuracy of the results.
tree_example_fitted_model = predict(tree_example_tree_gini, type= "class")

View(as.data.frame(tree_example_fitted_model))

#tree_example_fitted_model <- as.numeric(tree_example_fitted_model)
View(tree_example_fitted_model)

# Let's compare the results to the actual data.
preg_conf_matrix = table(tree_example_fitted_model, tree_example$PREGNANT)
preg_conf_matrix

table(tree_example_fitted_model)

#We can also just use the confusion matrix

library(caret)
confusionMatrix(as.factor(tree_example_fitted_model), 
                as.factor(tree_example$PREGNANT), 
                positive = "preg", 
                dnn=c("Prediction", "Actual"), 
                mode = "sens_spec")

table(tree_example$PREGNANT)

```

Hit Rate or True Classification Rate, Detection Rate and ROC
```{r}
# The error rate is defined as a classification of "Pregnant" when 
# this is not the case, and vice versa. It's the sum of all the
# values where a column contains the opposite value of the row.
sum(preg_conf_matrix[row(preg_conf_matrix)!= col(preg_conf_matrix)])
# 301


# The error rate divides this figure by the total number of data points
# for which the forecast is created.
sum(preg_conf_matrix)
# 2000

# Let's use these values in 1 calculation.
preg_error_rate = sum(preg_conf_matrix[row(preg_conf_matrix) != col(preg_conf_matrix)])/ sum(preg_conf_matrix)


paste0("Hit Rate/True Error Rate:", preg_error_rate * 100, "%")
# "Hit Rate/True Error Rate:15.05%"


#Detection Rate is the rate at which the algo detects the positive class in proportion to the entire classification A/(A+B+C+D) where A is poss poss

preg_conf_matrix

preg_conf_matrix[2,2]/sum(preg_conf_matrix)# 17.75%, want this to be higher but only so high it can go, in a perfect model for this date it would be:


preg_roc <- roc(tree_example$PREGNANT, as.numeric(tree_example_fitted_model), plot = TRUE) #Building the evaluation ROC and AUV using the predicted and original target variables 

preg_roc

plot(preg_roc)

#We can adjust using a if else statement and the predicted prob

tree_example_fitted_prob = predict(tree_example_tree_gini, type= "prob")
View(tree_example_fitted_prob)

#Let's 
roc(tree_example$PREGNANT, ifelse(tree_example_fitted_prob[,'not_preg'] >= .25,0,1), plot=TRUE)

```

#We can also prune the tree to make it less complex 
```{r}
set.seed(1)
tree_example_tree_cp2 = rpart(PREGNANT~.,#<- formula, response variable ~ predictors,"." means "use all other variables in data"
                           method = "class", #<- specify method, use "class" for tree
                           parms = list(split = "gini"),#<- method for choosing tree split
                           data = tree_example,#<- data used
                           control = rpart.control(maxdepth = 7)) #<- includes depth zero, the control for additional options (could use CP, 0.01 is the default)

?rpart.control

plotcp(tree_example_tree_cp2)

View(tree_example_tree_cp2)

rpart.plot(tree_example_tree_cp2, type =4, extra = 101)

cptable_ex_cp <- as.data.frame(tree_example_tree_cp2$cptable, )
View(cptable_ex_cp)

cptable_ex_cp$opt <- cptable_ex_cp$`rel error`+ cptable_ex_cp$xstd

View(cptable_ex_cp)

#Change the rpart.control and take a look at results.

dev.off()

```
