---
title: "In Class DT"
author: "Brian Wright"
date: "December 7, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Congrats! You just graduated UVA's MSDS program and got a job working at the 
Treasury Department. In partnership with Congress the Treasury has been 
ask to come up with innovative ways to create tax policy. In doing so they 
want to be able to reliably predict whether American workers make more than 
$50,000 a year.  

You would like to be able to explain the model to the mere mortals 
but need a fairly robust and flexible approach so you've 
chosen to use decision trees to get started. 

In doing so, similar to  great data scientists of the past 
you remembered the excellent education provided 
to you at UVA in a undergrad data science course and have outline 
20ish steps that will need to be undertaken to complete this task. As always, you will need to make sure to #comment your work heavily. 


 Footnotes: 
-	You can add or combine steps if needed
-	Also, remember to try several methods during evaluation and always be 
mindful of how the model will be used in practice.
- Make sure all your variables are the correct type (factor, character, etc.)


```{r}
url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data"

names <- c("age","workclass","fnlwgt","education","education-num","marital-status","occupation","relationship","race","sex","capital-gain","capital-loss","hours-per-week","native-country", "salary")

xx <- readr::read_csv(url,col_names=names)

View(xx)

```

```{r}
names <- c("age","workclass","fnlwgt","education","education-num","marital-status","occupation","relationship","race","sex","capital-gain","capital-loss","hours-per-week","native-country", "salary")
```

```{r}
#1 Load the data and ensure the labels are correct. Restate the question and comment on evaluation metrics that you might pay attention to.  
```

```{r}
#2 Ensure all the variables are classified correctly including the target variable and collapse factor variables as needed. 
```

```{r}
#3 Check for missing variables and correct as needed.  
```

```{r}
#4 Guess what, you don't need to standardize the data,because DTs don't require this to be done, they make local greedy decisions...keeps getting easier, go to the next step
```

```{r}
#5 Determine the baserate or prevalence for the classifier, what does this number mean? (Need to re-code the target)  
```

```{r}
#6 Split your data into test, tune, and train. (80/10/10)
```

```{r}
#7 Build your model using the training data, rpart2, and repeated cross validation as reviewed in class with the caret package.
```

```{r}
#8 View the results, comment on how the model performed and which variables appear to be contributing the most (variable importance)  
```

```{r}
#9 Plot the output of the model to see the tree visually, using rpart.plot 
```

```{r}
#10 Use the tune set and the predict function with your model to the 
# predict the target variable, making sure to produce probabilities.
```

```{r}
#11 Use the the confusion matrix function on your predictions to check a variety of metrics and comment on the metric that might be best for this type of analysis given your question.  
```

```{r}
#12 With the percentages you generated in step 10,  select several different threshold levels using the threshold function we created and interpret the results. What patterns do you notice, did the evaluation metrics change? 
```

```{r}
#13 Based on your understanding of the model and data adjust the hyper-parameter via the built in train control function in caret or build and try new features, does the model quality improve or not? If so how and why, if not, why not?
```

```{r}
#14 Once you are confident that your model is not improving, via changes implemented on the training set and evaluated on the the tune set, predict with the test set and report final evaluation of the model. Discuss the output in comparison with the previous evaluations.  
```

```{r}
#17 Summarize what you learned along the way and make recommendations on how this could be used moving forward, being careful not to over promise. 
```

```{r}
#18 What was the most interesting or hardest part of this process and what questions do you still have? 
```


