---
title: "KNN_lab_2"
author: "Brian Wright"
date: "8/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


I think this will work need to drop a bunch of columns but not too bad. Try to predict something to do with % graduation. 

Instructions:
The data is messy and you have a degrees of freedom problem, as in, you have too
many features. Using the readme and your groups intuition to reduce the number of variables down to around 20ish. 

Engineer the target variable using the grad_150_percentile into a binary variable. Use the distribution looking for natural breaks points or other intuitions from your group. You can also search for publications that elude to 
what a high quality 6 year graduation rate might be.


```{r}
url <- "https://query.data.world/s/yd5wiazzlj7aahmn4x37y7zq5pyh2h"

grad_data <- read_csv(url)

View(grad_data)

# readme for the dataset - https://data.world/databeats/college-completion/workspace/file?filename=README.txt

#Pretty large number of features, need to trim this down
dim(grad_data_2)

#This will allow us to pull column numbers for sub-setting  
column_index <- tibble(colnames(grad_data))

x <- 40:56 #create a list so we don't have to type 16 numbers out

#Most of these columns have a good number of missing values or are not useful.  
grad_data_1 <- grad_data[ ,c(-28,-10,-11,-12,-x,-29,-37,-61,-57)]

column_index <- tibble(colnames(grad_data_1))

View(column_index)

View(grad_data_2)

grad_data_2 <- grad_data_1[ ,c(-7,-12,-13,-15,-16,-19,-22,-24,-30,-33,-34,-35,-36)]



```

