---
title: "reinforce"
author: "Anna Stein"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  
```

```{r}
url <- "https://raw.githubusercontent.com/ahs2aew/DS-3001/main/week-11-reinforcement-lab/data-frame.csv"
reinforce_data <- read_csv(url)

View(reinforce_data)
```

```{r}
ggplot(reinforce_data,aes(x=cora,y=corc))+geom_point()
```
cora: how like the first deck the row is
corc: how like the other deck (very different) the row is 



what would clusters be based on? 
- maybe 3 clusters
     - high cora, low/medium corc
     - mid cora and mid corc
     - mid cora and higher corc 
     
```{r}
clust_data_reinforce = data[,c("cora", "corc")]
clust_data_reinforce # these are the columns we want to look at for the clustering 


set.seed(1)
kmeans_obj_reinforce = kmeans(clust_data_reinforce, centers = 2, 
                        algorithm = "Lloyd") 
kmeans_obj_reinforce
```
With 2 clusters, the variance explained is 45.5%. This is pretty low. We should see if we can get this higher if we have more clusters. 


```{r}
cor_clusters = as.factor(kmeans_obj_reinforce$cluster) # the cluster will be the factor: either 1 or 2
View(as.data.frame(cor_clusters))

# this tells us which cluster each point belongs to 
```


Visualize the clusters: 
```{r}
reinforce_plot= ggplot(reinforce_data, aes(x = cora, 
                            y = corc,
                            shape = cor_clusters)) + 
  geom_point(size = 6) +
  ggtitle("Title") +
  xlab("Cora") +
  ylab("Corc") +
  scale_shape_manual(name = "Cluster", 
                     labels = c("Cluster 1", "Cluster 2"),
                     values = c("1", "2")) +
  theme_light()
```
This looks weird - add more clusters?

```{r}
# save plot
ggsave(
  filename = 'Reinforce Plot',
  plot = last_plot(),  
  device = "png",
  path = NULL,
  scale = 1,
  width = NA,
  height = NA,
  units = c("in", "cm", "mm", "px"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
```
```{r}
getwd()

```



