---
title: "dplyr_Lab_Assignment"
author: "Brian Wright"
date: "2/9/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


################## AND NOW THE EXCITEMENT BEGINS! DPLYR AND TIDYVERSE PRACTICE 

## As a reminder five key dplyr functions that allow you to solve "the vast majority of your data-manipulation
## challenges" are:

## 1) filter() -- Pick observations by their values
## 2) arrage() -- Reorder the rows
## 3) select() -- Pick variables by their names
## 4) mutate() -- Create new variables with functions of existing variables
## 5) summarize() -- Collapse many values down to a single summary
## Bonus: group_by -- we can use this in combination with any of the five to change the scope of
## each function from operating on the entire dataset to operating on it 
group-by-group.


For this assignment you can use any or all of the datasets I 
provided or choose three of your own!  I would like you to compose 
one question for each dataset (or dataset of your choosing) that requires at a minimum three functions listed above to answer. Do not repeat function sequences or questions.  Information or the actual data for the datasets can be found through the links below. In summary for each dataset produce four parts: 1)Question, 2)psuedocode that answers the question, 3)single line transactions using dplyr functions, 4) pipe the individual lines together and create a function. 

Also you can work with your groups for coding tips/advise or work through similar programming issues but everyone must have their own set of questions and results. 

# any dataset in the data file is fine or 

# nf2008_fg: http://users.stat.ufl.edu/~winner/data/nfl2008_fga.csv

# red_wine_quality: https://data.world/uci/wine-quality


# Lets walk through a example. Below is a dataset on students admitted to a graduate program in Data Science (actually don't remember where this data came from). 

## Are there differences in the admitted number of students on average by GPA and school rank?

  >A easy way to handle this is to identify function calls starting at the end of the question and working backwards. 

## Pseudocode 
  - Group by school rank
  - Calculate GPA Average
  - Count total admitted 
  - Calculate % Admitted
  - Arrange in descending order by Mean GPA


Loading and preparing the data 
```{r}
admit_2 <- read_csv("~/git_3001/DS-3001/data/LogReg.csv")

#creating a quick list to use as a index
names <- c("admit","rank")

names
#changing those two variables to factors
admit_2[,names] <- lapply(admit_2[,names],factor)

str(admit_2)

```

Single line version 
```{r}
#Group
View(admit_2)
x1 <- group_by(admit_2,rank)
#Calculate additional variables
x2 <- summarize(x1,admit=sum(admit == 1),count=n(),mean(gpa))
x2$ave <- x2$admit/x2$count
#Arrange by GPA
x3 <- arrange(x2,desc(`rank`))
print(x3)
## Are there differences in the admitted number of students on average by GPA and school rank?

```

## Piping it all together
```{r}
#creating a dplyr version
(admit_1 <- group_by(admit_2, rank)%>%
 summarize(admit=sum(admit == 1),count=n(),mean(gpa))%>%
 mutate(ave_admit=admit/count)%>%
 arrange(desc(`rank`)))

```


## Now just turn it into a function, so if you needed to repeat this data request on future datasets it would be simple. 
```{r}
ave_admit <- function(x){
  xx <- group_by(x, rank)%>%
  summarize(admit=sum(admit == 1),count=n(),mean(gpa))%>%
  mutate(ave_admit=admit/count)%>%  
  arrange(desc(`mean(gpa)`))
  print(xx)
}

ave_admit(admit_2)

```





















































```{r}

beaches <- xx %>% 
  group_by(season) %>% 
  sum(as.numeric(rainfall))

sum(xx$rainfall, na.rm = TRUE)
is.na(xx$rainfall)


head(xx)

```

